---
title: "KNN and Naive Bay"
author: "Yuxuan Wang"
date: "2023/4/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(fastDummies)
library(tidyverse)
library(caret)
library(MASS)
# read data ------------------------
math = read.csv("student-mat.csv", sep = ';', stringsAsFactors = T)
str(math)
por = read.csv("student-por.csv", sep = ";", stringsAsFactors = T)
str(por)

#set dummies
math = fastDummies::dummy_columns(math, remove_first_dummy = T, remove_selected_columns = T)
por = fastDummies::dummy_columns(por, remove_first_dummy = T, remove_selected_columns = T)


predictors = setdiff(colnames(math), c("G1", "G2", "G3"))

math$grade.cat = cut(math$G1, breaks=c(-1, 9, 11, 13, 15, 20),
                     labels=c('F', 'D', 'C', 'B', 'A'))
por$grade.cat = cut(por$G1, breaks=c(-1, 9, 11, 13, 15, 20),
                    labels=c('F', 'D', 'C', 'B', 'A'))
por$grade.con = por$G1
math$grade.con = math$G1


# split data -------------------------
set.seed(114514)
train_idx.math = sample(1:nrow(math), floor(nrow(math))*0.8, replace = F)
train_idx.por = sample(1:nrow(por), floor(nrow(por))*0.8, replace = F)

math.train = math[train_idx.math,]
math.test = math[-train_idx.math,]

por.train = por[train_idx.por,]
por.test = por[-train_idx.por,]

```
```{r}
par(mfrow=c(2,2))
hist(as.numeric(math.train$grade.cat),breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5),right=TRUE,freq=TRUE,main="Math Train")
hist(as.numeric(math.test$grade.cat),breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5),right=TRUE,freq=TRUE,main="Math Test")
hist(as.numeric(por.train$grade.cat),breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5),right=TRUE,freq=TRUE,main="Por Train")
hist(as.numeric(por.test$grade.cat),breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5),right=TRUE,freq=TRUE,main="Por Test")
```
```{r}
#math$grade.cat = cut(math$G1, breaks=c(-1, 9, 20),labels=c("Low","High"))
#por$grade.cat = cut(por$G1, breaks=c(-1, 9, 20),labels=c("Low","High"))
set.seed(114514)
train_idx.math = sample(1:nrow(math), floor(nrow(math))*0.8, replace = F)
train_idx.por = sample(1:nrow(por), floor(nrow(por))*0.8, replace = F)

math.train = math[train_idx.math,]
math.test = math[-train_idx.math,]

por.train = por[train_idx.por,]
por.test = por[-train_idx.por,]
preproc.param <- math.train %>%
  preProcess(method = c("center", "scale"))
math.train <- preproc.param %>% predict(math.train)
math.test <- preproc.param %>% predict(math.test)
preproc.param <- por.train %>%
  preProcess(method = c("center", "scale"))
por.train <- preproc.param %>% predict(por.train)
por.test <- preproc.param %>% predict(por.test)

library(e1071)
classification.formula = as.formula(paste("grade.cat~", paste(predictors, collapse = "+"), sep = ''))
regression.formula = as.formula(paste("grade.con~", paste(predictors, collapse = "+"), sep = ''))
```

## KNN Regression
```{r}
library(plotrix)
library(FNN)
library(e1071)
## Math-----------------

ctrl <- trainControl(method = "cv", number = 10)

knn.model.math <- train(regression.formula, data = math.train, preProcess = c("center", "scale"), method = "knn", trControl = ctrl, tuneLength = 20)
plot(knn.model.math)

bestK.math <- knn.model.math$bestTune$k
bestK.math
Knn.for.math <- knn.reg(train = math.train[,-c(14,15,16,43)], test = math.test[,-c(14,15,16,43)], y = math.train$grade.con, k = bestK.math)

#MSE
mean((math.test$grade.con - Knn.for.math$pred)^2)

Knn.for.math <- knn.reg(train = math.train[,-c(14,15,16,43)], test = math.train[,-c(14,15,16,43)], y = math.train$grade.con, k = bestK.math)

#MSE
mean((math.train$grade.con - Knn.for.math$pred)^2)
## Por---------------
# For Portuguese

knn.model.por <- train(regression.formula, data = por.train, preProcess = c("center", "scale"), method = "knn", trControl = ctrl, tuneLength = 20)
plot(knn.model.por)
bestK.por <- knn.model.por$bestTune$k
bestK.por

Knn.for.por <- knn.reg(train = por.train[,-c(14,15,16,43)], test = por.test[,-c(14,15,16,43)], y = por.train$grade.con, k = bestK.por)

# MSE
mean((por.test$grade.con - Knn.for.por$pred)^2)

Knn.for.por <- knn.reg(train = por.train[,-c(14,15,16,43)], test = por.train[,-c(14,15,16,43)], y = por.train$grade.con, k = bestK.por)

# MSE
mean((por.train$grade.con - Knn.for.por$pred)^2)
```

## KNN Classification
```{r}
## math----------
library(class)
set.seed(114514)
ctrl <- trainControl(method = "cv", number = 10)

knn.model.math <- train(classification.formula, data = math.train, preProcess = c("center", "scale"), method = "knn", trControl = ctrl, tuneLength = 20)
plot(knn.model.math)
bestK.math <- knn.model.math$bestTune$k
bestK.math
math.train$grade.cat<-as.numeric(math.train$grade.cat)
math.test$grade.cat<-as.numeric(math.test$grade.cat)
Knn.for.math <- knn(train = math.train[,-c(14,15,16,44)], test = math.test[,-c(14,15,16,44)], cl = math.train$grade.cat, k = bestK.math)
#Misclassification Error
1-mean(math.test$grade.cat==Knn.for.math)
Knn.for.math <- knn(train = math.train[,-c(14,15,16,44)], test = math.train[,-c(14,15,16,44)], cl = math.train$grade.cat, k = bestK.math)
#Misclassification Error
1-mean(math.train$grade.cat==Knn.for.math)
## por----------
set.seed(114514)
ctrl <- trainControl(method = "cv", number = 10)

knn.model.por <- train(classification.formula, data = por.train, preProcess = c("center", "scale"), method = "knn", trControl = ctrl, tuneLength = 20)
plot(knn.model.por)
bestK.por <- knn.model.por$bestTune$k
bestK.por
por.train$grade.cat<-as.numeric(por.train$grade.cat)
por.test$grade.cat<-as.numeric(por.test$grade.cat)
Knn.for.por <- knn(train = por.train[,-c(14,15,16,44)], test = por.test[,-c(14,15,16,44)], cl = por.train$grade.cat, k = bestK.por)
#Misclassification Error
1-mean(por.test$grade.cat==Knn.for.por)
Knn.for.por <- knn(train = por.train[,-c(14,15,16,44)], test = por.train[,-c(14,15,16,44)], cl = por.train$grade.cat, k = bestK.por)
#Misclassification Error
1-mean(por.train$grade.cat==Knn.for.por)
```

## Naive Bayes Classifier

library(naivebayes)
# For Math

tune_grid <- expand.grid(
  laplace = seq(0.005, 0.05, by=0.005),
  usekernel = FALSE,
  adjust = c(TRUE,FALSE))

nbc.model.math <- train(classification.formula,data = math.train,method = "naive_bayes",trControl = ctrl, tuneGrid = tune_grid)

plot(nbc.model.math)



nbc.model.math$bestTune

# For Portuguese

nbc.model.por <- train(classification.formula,data = por.train,method = "naive_bayes",trControl = ctrl, tuneGrid = tune_grid)

plot(nbc.model.por)
nbc.model.por$bestTune


